A Proposal for an Intuitionistic Fuzzy Inference System

### 2016-01-08 paper.tex
Done: Experiments, Results, Conclusion, Future Work

\section{Proposed Method}
%(Mario)Is a Method? Extension? Implementation?
%% (A): No lo había pensado en realidad. Es que siempre le llamo
%% Proposed Method jajajaaj. Aunque creooo que sí es un método. Es un
%% método para hacer un sistema de inferencia difusa con conjuntos
%% difusos intuicionistas
The method, for now, consists on the procedures and elements
necessary to construct a basic Intuitionistic Fuzzy Inference System
(IFIS). %(Mario)Por que basic?
%% (A): Basic porque no añado que puedas poner OR en las reglas, no
%% pongo otros métodos de defuzzificación, no pongo más funciones de
%% membresía (triangulares, etc.). Explicaré eso (y)
%(Mario)Tal vez enumerarlas? cuales procedures son?
%% (A): Los procedures son los que están en el diagramita, y las
%% explico abajo. Aun así las enumero?
An implementation was completed in order to make comparisons against
classic Fuzzy Inference Systems (FISs), such as Type-1 FIS, Interval
Type-2 and General Type-2 FIS, and prove that
the design was viable and able to contribute to better results than
the classic FISs.
%(Mario)Es el proposito del diseño comparar? Creo que puedes decir mejor algo como:
%In order to assess the advantage of our proposal in terms of X
%a comparassion with  Type-1 FIS, Interval Type-2 and General Type-2 FIS,
%is presented.

@@

In many works, intuitionistic fuzzy sets (IFSs) are constructed manualy,
by interviewing experts who assign memberships and
non-memberships to each of the values of a set (such as in %to each of the members of a set?
\cite{davarzani2013novel}). This work proposes a convenient way of
creating an IFS, imitating many systems that use Gaussian
membership functions (MFs) to facilitate an expert the representation of a
fuzzy interpretation. An example of an IFS is depicted in
Fig. \ref{fs-as-ifs}. This IFS represents a classic fuzzy set, as it
satisfies (\ref{ifs-form}) and (\ref{fs-as-ifs-if}), by representing
the MF as the red line, and the non-membership
function (NMF) as the blue line. In order to contruct the NMF, one has
to rotate a normal Gaussian MF by 180˚, and this can be seen as the
complement of the fuzzy set corresponding to a Gaussian MF.
%(Mario)Al parecer lo que se está presentando es una manera automatizada de 
% hacer IFSs? Si  es así, es más facil presentarlo
%% Si se hacen automatizadamente para poder compararlo con el tipo-1 y
%% tipo-2, pero no es el objetivo del trabajo.

@@

Two convenient ways of modifying a NMF and also
satisfying (\ref{intuitionistic-interval}) are the following: 1) setting
the mean and the standard deviation of the Gaussian functions to be
the same, and only vary the $height$ of the NMF,
as is depicted in Fig. \ref{ifs}; and 2) setting different means and
standard deviations of the Gaussian membership and non-membership
functions, while keeping the $height$ of the MF equal
to $h_{mf} = 1 - h_{nmf}$, as this ensures that
(\ref{intuitionistic-interval}) is satisfied. An example of the latter
case is shown in Fig. \ref{ifs-diff-mu-sd}. To clarify, the $height$
of the MF ($h_{mf}$) of an IFS is equal to $max(\mu(x))$, and the
$height$ of the NMF ($h_{nmf}$) of an IFS is equal to $max(\nu(x))$.

@@

The first step of the inference process in a classic FIS (a flow chart
for the process in an IFIS is given in Fig. \ref{flow-chart}) is to
determine what grade of membership corresponds to a given input, given a set of
%(Mario)determine what particular membership function corresponds to a 
% given input, given a set of 
%% (A): Es grade of membership, perdón. De hecho hay ese error en
%% otras partes, tengo que cambiarlo.
MFs. This is straightforward in a classic FIS, as the
membership to each input is already established in the membership
fuzzy set. For an IFIS, this work proposes a method to determine a
membership given a MF and a NMF,
which is described in (\ref{imembership}), and receives the name of
$i\mu$ (intuitionistic membership). %(Mario)Creo que este podría ser
                                %el titulo
%% (A): Para poder hacer un FIS intuicionista, necesité el
%% intuitionistic membership y el intuitionistic center of area. Cómo
%% propone el título?

@@

A comparison of a MF (red line) against an intuitionistic
membership function (blueline) is shown in Fig.
\ref{if-membership}. The parameters for both the MF and NMF are $\mu =
50$ and $SD = 15$, but the $height$ for the NMF is set at $0.5$. This
results in an intuitionistic MF that appears to have the same mean,
but a smaller standard deviation. A more elaborated example is shown
in Fig. \ref{if-membership-drastic}, where the parameters of the MF
are $\mu = 50$, $SD = 15$ and $h = 0.8$, and the parameters of the NMF
are $\mu = 30$, $SD = 20$ and $h = 0.2$. As can be noted, the
indeterminacy ($\pi$) is high in this example, which results in a
smaller $SD$ for the intuitionistic MF. Moreover, the difference in
means and standard deviations generates a change of the mean for the
intuitionistic MF.

@@

The next step in the inference process is to perform an $\alpha$-cut % (Mario) La del FIS? aclaralo
on the associated consequents to the antecedents that are activated by
the inputs. In a classic FIS, this process is straightforward, as the
memeberships in a fuzzy set representing a consequent are ``cut'' by
following (\ref{alpha-cut}) for every element in a fuzzy set
representing a consequent. In an intuitionistic MF, one can follow
the same procedure in order to obtain the $\alpha$-cut in the MF part,
but to obtain the $\alpha$-cut in the NMF, (\ref{nmf-alpha-cut}) must
be followed. In (\ref{nmf-alpha-cut}), the counterintuitive part is
that one has to find the corresponding $\nu$ for a given
$\mu_{\alpha}$. An example of an $\alpha$-cut, with $\mu_{\alpha} =
0.6$, is depicted in Fig. \ref{alpha-cut-example}.

@@

After performing every necessary $\alpha$-cut, one has to apply the
union operator over every resulting fuzzy set, in order to aggregate
the weights of all the consequents. The union operator is defined by
K. Atanassov \cite{atanassov2013intuitionistic}, for every two IFSs A
and B, as in (\ref{union-operator}). An example of an union of three
IFSs is shown in Fig. (\ref{ifs-union}).

@@

Fig. \ref{if-coa-vs-coa} shows the plot of the defuzzification of
several IFSs with $CoA$ (red line) and $iCoA$ (blue line), with means set
from 0 to 100 for the MF, and -20 to 80 for the NMF, and standard
deviations of 15 for the MF, and 20 for the NMF.

@@

The Mackey-Glass time series benchmark was used in order to compare
the performances of each of the FISs, and a Genetic Algorithm (GA) was
used to optimize each of the FISs, so sub-optimal versions of each of
the systems were used for the comparison.

@@

@@

Regarding the parameters of the GA, every FIS is optimized for 20
generations, using 100 individuals as the initial population. The
roulette wheel method is used as the selection operator. Five
children are generated each generation and replace the five least
fitted individuals in the population. For the mutation operator, each
individual in the population has a chance of 1\% of suffering an
alteration, which consists on replacing a list of tuples of arbitrary
length from the individual with a

@@

\section{Results}

The results of the experiment along with the code can be found on the 
following public Github repository: https://github.com/amherag/wcci-2016-a

\begin{table}[!t]
  \renewcommand{\arraystretch}{1.3}
  \caption{Hypothesis Tests for the Training Stage}
  \label{hypothesis-tests-training}
  \centering
  \begin{tabular}{|c|c|c|c|c|c|}
    \hline
    Method & $\mu_{TR}$ & $SD_{TR}$ & $n$ & t-Value\\
    \hline
    T1 iFIS & 89.50 & 25.36 & 60 & \\
    \hline
    T1 FIS & 108.05 & 23.47 & 60 & -4.1584 \\
    \hline
    IT2 FIS (\(\mu\)) & 112.56 & 29.16 & 100 & -5.2597 \\
    \hline
    IT2 FIS (SD) & 102.31 & 26.08 & 300 & -3.5548 \\
    \hline
  \end{tabular}
\end{table}

\begin{table}[!t]
  \renewcommand{\arraystretch}{1.3}
  \caption{Hypothesis Tests for the Testing Stage}
  \label{hypothesis-tests}
  \centering
  \begin{tabular}{|c|c|c|c|c|c|}
    \hline
    Method & $\mu_{TS}$ & $SD_{TS}$ & $n$ & t-Value\\
    \hline
    T1-iFIS & 160.99 & 65.70 & 60 & \\
    \hline
    T1-FIS & 194.97 & 71.97 & 60 & -2.7010 \\
    \hline
    IT2-FIS (\(\mu\)) & 192.31 & 105.75 & 100 & -2.3104 \\
    \hline
    IT2-FIS (SD) & 178.63 & 99.70 & 300 & -1.7209 \\
    \hline
  \end{tabular}
\end{table}

@@

This work mainly focuses on proving that the design is viable, but it
also proves that at least in some cases the intuitionistic fis can
out-perform the t1 and t2.

@@

We want to create a toolbox such as \cite{castro2007interval}
We want to experiment with other data sets.
We want to speed up the algorithm.
We want to experiment with generalized type-2.

>>>>

\section{Proposed Method}
\label{proposed-method}
%(Mario)Is a Method? Extension? Implementation?
%% (A): No lo había pensado en realidad. Es que siempre le llamo
%% Proposed Method jajajaaj. Aunque creooo que sí es un método. Es un
%% método para hacer un sistema de inferencia difusa con conjuntos
%% difusos intuicionistas
The method, for now, consists on the procedures and elements
necessary to construct a basic Intuitionistic Fuzzy Inference System
(IFIS). %(Mario)Por que basic?
%% (A): Basic porque no añado que puedas poner OR en las reglas, no
%% pongo otros métodos de defuzzificación, no pongo más funciones de
%% membresía (triangulares, etc.). Explicaré eso (y)
%(Mario)Tal vez enumerarlas? cuales procedures son?
%% (A): Los procedures son los que están en el diagramita, y las
%% explico abajo. Aun así las enumero?
An implementation was completed in order to make comparisons against
classic Fuzzy Inference Systems (FISs), such as Type-1 FIS, Interval
Type-2 and General Type-2 FIS, and prove that
the design was viable and able to contribute to better results than
the classic FISs.
%(Mario)Es el proposito del diseño comparar? Creo que puedes decir mejor algo como:
%In order to assess the advantage of our proposal in terms of X
%a comparassion with  Type-1 FIS, Interval Type-2 and General Type-2 FIS,
%is presented.

@@

In many works, intuitionistic fuzzy sets (IFSs) are constructed manualy,
by interviewing experts who assign memberships and
non-memberships to each of the values of a set (such as in %to each of the members of a set?
\cite{davarzani2013novel}). This work proposes a convenient way of
creating an IFS, imitating many systems that use Gaussian
membership functions (MFs) to facilitate an expert the representation of a
fuzzy interpretation. An example of an IFS is depicted in
Figure \ref{fs-as-ifs}. This IFS represents a classic fuzzy set, as it
satisfies (\ref{ifs-form}) and (\ref{fs-as-ifs-if}), by representing
the MF as the red line, and the non-membership
function (NMF) as the blue line. In order to contruct the NMF, one has
to rotate a normal Gaussian MF by 180˚, and this can be seen as the
complement of the fuzzy set corresponding to a Gaussian MF.
%(Mario)Al parecer lo que se está presentando es una manera automatizada de 
% hacer IFSs? Si  es así, es más facil presentarlo
%% Si se hacen automatizadamente para poder compararlo con el tipo-1 y
%% tipo-2, pero no es el objetivo del trabajo.

@@

Two convenient ways of modifying a NMF and also
satisfying (\ref{intuitionistic-interval}) are the following: 1) setting
the mean and the standard deviation of the Gaussian functions to be
the same, and only vary the $height$ of the NMF,
as is depicted in Figure \ref{ifs}; and 2) setting different means and
standard deviations of the Gaussian membership and non-membership
functions, while keeping the $height$ of the MF equal
to $h_{mf} = 1 - h_{nmf}$, as this ensures that
(\ref{intuitionistic-interval}) is satisfied. An example of the latter
case is shown in Figure \ref{ifs-diff-mu-sd}. To clarify, the $height$
of the MF ($h_{mf}$) of an IFS is equal to $max(\mu(x))$, and the
$height$ of the NMF ($h_{nmf}$) of an IFS is equal to $max(\nu(x))$.

@@

The first step of the inference process in a classic FIS (a flow chart
for the process in an IFIS is given in Figure \ref{flow-chart}) is to
determine what grade of membership corresponds to a given input, given a set of
%(Mario)determine what particular membership function corresponds to a 
% given input, given a set of 
%% (A): Es grade of membership, perdón. De hecho hay ese error en
%% otras partes, tengo que cambiarlo.
MFs. This is straightforward in a classic FIS, as the
membership to each input is already established in the membership
fuzzy set. For an IFIS, this work proposes a method to determine a
membership given a MF and a NMF,
which is described in (\ref{imembership}), and receives the name of
$i\mu$ (intuitionistic membership). %(Mario)Creo que este podría ser
                                %el titulo
%% (A): Para poder hacer un FIS intuicionista, necesité el
%% intuitionistic membership y el intuitionistic center of area. Cómo
%% propone el título?

@@

A comparison of a MF (red line) against an intuitionistic
membership function (blueline) is shown in Figure
\ref{if-membership}. The parameters for both the MF and NMF are $\mu =
50$ and $SD = 15$, but the $height$ for the NMF is set at $0.5$. This
results in an intuitionistic MF that appears to have the same mean,
but a smaller standard deviation. A more elaborated example is shown
in Figure \ref{if-membership-drastic}, where the parameters of the MF
are $\mu = 50$, $SD = 15$ and $h = 0.8$, and the parameters of the NMF
are $\mu = 30$, $SD = 20$ and $h = 0.2$. As can be noted, the
indeterminacy ($\pi$) is high in this example, which results in a
smaller $SD$ for the intuitionistic MF. Moreover, the difference in
means and standard deviations generates a change of the mean for the
intuitionistic MF.

@@

The next step in the inference process is to perform an $\alpha$-cut % (Mario) La del FIS? aclaralo
on the associated consequents to the antecedents that are activated by
the inputs. In a classic FIS, this process is straightforward, as the
memeberships in a fuzzy set representing a consequent are ``cut'' by
following (\ref{alpha-cut}) for every element in a fuzzy set
representing a consequent. In an intuitionistic MF, one can follow
the same procedure in order to obtain the $\alpha$-cut in the MF part,
but to obtain the $\alpha$-cut in the NMF, (\ref{nmf-alpha-cut}) must
be followed. In (\ref{nmf-alpha-cut}), the counterintuitive part is
that one has to find the corresponding $\nu$ for a given
$\mu_{\alpha}$. An example of an $\alpha$-cut, with $\mu_{\alpha} =
0.6$, is depicted in Figure \ref{alpha-cut-example}.

@@

After performing every necessary $\alpha$-cut, one has to apply the
union operator over every resulting fuzzy set, in order to aggregate
the weights of all the consequents. The union operator is defined by
K. Atanassov \cite{atanassov2013intuitionistic}, for every two IFSs A
and B, as in (\ref{union-operator}). An example of an union of three
IFSs is shown in Figure (\ref{ifs-union}).

@@

Figure \ref{if-coa-vs-coa} shows the plot of the defuzzification of
several IFSs with $CoA$ (red line) and $iCoA$ (blue line), with means set
from 0 to 100 for the MF, and -20 to 80 for the NMF, and standard
deviations of 15 for the MF, and 20 for the NMF.

@@

The Mackey-Glass time series benchmark was used in order to compare
the performances of each of the FISs regarding their time series
forecasting capabilities, and a Genetic Algorithm (GA) was
used to optimize each of the FISs, so sub-optimal versions of each of
the systems were used for the comparison.

@@

The dataset consists of 100 data points for training, and 50 data
points for testing. These data points were scaled, such that the
resulting data points range from 0 to 100. The scaling process was
performed by applying (\ref{scaling}) over each of the data points in
the union of the training and testing subsets. The reason behind this
preprocessing is that the implementation of the IFIS works with inputs
ranging from 0 to 100, but the Mackey-Glass time series' data points
range from 0 to about 1.4. This does not alter the measurement of the
error produced by each of the FISs, as the outputs from the FISs are
also scaled from 0 to 100.

\begin{equation}
  \label{scaling}
  S(p) = 100 \frac{p - min(P)}{max(P) - min(P)}
\end{equation}

@@

Regarding the parameters of the GA, every FIS is optimized for 20
generations, using 100 individuals as the initial population. The
roulette wheel method is used as the selection operator. Five
children are generated each generation and replace the five least
fitted individuals in the population. For the mutation operator, each
individual in the population has a chance of 1\% of suffering an
alteration, which consists on replacing a list of tuples of arbitrary
length from the affected individual with a list of tuples of the same
length from another random individual in the population. Finally, the
crossover operator takes random genetic material from two individuals
and creates a child of the same length as its parents. The genetic
material cannot be repeated, can be of different proportions from each
of the parents, and the original order isn't preserved. For example,
given the individuals $A = <1, 2, 3, 4>$ and $B = <5, 6, 7, 8>$, a
possible child is $C = <8, 1, 7, 5>$. This operator was developed
because it proved by trial and error that it created more diversity in
the children than other operators such as one-point or two-point
crossover.

To obtain the performance of a particular individual, a FIS is created
using the parameters represented by its genetic material. The created
FIS is then evaluated with each of the inputs in the training dataset
(100 sets of four data points: $t_{0}$, $t_{-6}$, $t_{-12}$, and
$t_{-18}$). The outputs (i.e. the 100 predicted values) for each of
the inputs are then scaled from 0 to 100, as was mentioned before, so
a Mean Squared Error (MSE) can be obtained by comparing the predicted values
to the scaled Mackey-Glass' points ($p_{t+1}$). Individuals with lower
MSEs are said to be performing better.

Finally, the better performing individual after the 20 generations in
the GA is evaluated using the training and testing datasets, and the
resulting MSEs are used as the result of one experiment. In total, 60
experiments were performed for the T1-IFIS, 60 for the T1-FIS, 100 for
the IT2-FIS with uncertain mean, and 300 for the IT2-FIS with
uncertain standard deviation.

@@

\section{Results}
\label{results}

The full results (CSV files containing the MSE for each of the
experiments) along with the Clojure source code used to obtain the
data presented in this section can be found on the following public
Github repository: https://github.com/amherag/wcci-2016-a.

Initially, 60 experiments were performed for every FIS, but that
number of experiments didn't provide enough statistical data for the
IT2-FISs in order to make a conclusion. For the IT2-FIS with uncertain
mean, 100 experiments were needed to provide enough statistical data
to conclude what FIS performed better, and for the IT2-FIS with
uncertain standard deviation, 300 experiments were needed. To clarify,
all of the comparisons were against the T1-IFIS.

Table \ref{hypothesis-tests-training} summarizes the results obtained
for each of the FISs in the training stage, as well as the t-value
obtained when calculating a Student's t-test. The same structure is
followed in Table \ref{hypothesis-tests-testing}, which shows the
results obtained for each of the FISs in the testing stage.

\begin{table}[!t]
  \renewcommand{\arraystretch}{1.3}
  \caption{Hypothesis Tests for the Training Stage}
  \label{hypothesis-tests-training}
  \centering
  \begin{tabular}{|c|c|c|c|c|c|}
    \hline
    Method & $\mu_{TR}$ & $SD_{TR}$ & $n$ & t-Value\\
    \hline
    T1-IFIS & 89.50 & 25.36 & 60 & \\
    \hline
    T1-FIS & 108.05 & 23.47 & 60 & -4.1584 \\
    \hline
    IT2-FIS (\(\mu\)) & 112.56 & 29.16 & 100 & -5.2597 \\
    \hline
    IT2-FIS (SD) & 102.31 & 26.08 & 300 & -3.5548 \\
    \hline
  \end{tabular}
\end{table}

\begin{table}[!t]
  \renewcommand{\arraystretch}{1.3}
  \caption{Hypothesis Tests for the Testing Stage}
  \label{hypothesis-tests-testing}
  \centering
  \begin{tabular}{|c|c|c|c|c|c|}
    \hline
    Method & $\mu_{TS}$ & $SD_{TS}$ & $n$ & t-Value\\
    \hline
    T1-IFIS & 160.99 & 65.70 & 60 & \\
    \hline
    T1-FIS & 194.97 & 71.97 & 60 & -2.7010 \\
    \hline
    IT2-FIS (\(\mu\)) & 192.31 & 105.75 & 100 & -2.3104 \\
    \hline
    IT2-FIS (SD) & 178.63 & 99.70 & 300 & -1.7209 \\
    \hline
  \end{tabular}
\end{table}

@@

As the results in Section \ref{results} demonstrate, the T1-IFIS
outperforms the T1-FIS, IT2-FIS with uncertain mean, and IT2-FIS with
uncertain standard deviation. These results sustain the claim that the
intuitionistic grade of membership ($i\mu$), and the intuitionistic center
of area ($iCoA$), proposed in Section \ref{proposed-method} allows the
creation of a T1-IFIS that can perform better than a classic T1-FIS
and an IT2-FIS.

Several derivative works can result from the ideas presented in this paper. For
example, the proposed T1-IFIS can be tested using several
benchmarks, other than the Mackey-Glass time series. Also, this work
should inspire the creation of Type-2 Intuitionistic Fuzzy Inference
Systems (both the interval and general types).

@@

The current implementation of the proposed method lacks a Graphical User
Interface (GUI) where a research can conveniently design an IFIS, such as
Matlab's Type-1 Fuzzy Logic toolbox, or the Type-2 Fuzzy Logic toolbox
developed by J. R. Castro $et al.$ \cite{castro2007interval}. The
development of a GUI is planned as a next step, so researchers without
a background in the programming langague Clojure can use the
implementation for their works.

In theory, an implementation of the proposed method should be faster
than T2-FISs, and slower than T1-FISs (although this difference should
be negligible), but the current implementation performs worse, in
terms of time, than a general T2-FIS. This has to be solved in order
to speed up the experiments conducted, and to encourage other
researchers to use the implementation.
